\textit{Supplementary materials for}: Text embedding models yield high-resolution insights into conceptual knowledge from short multiple-choice quizzes

### 2025-03-02 paper/supplement.tex

In our study, we relate the conceptual contents of course lectures and quiz 
questions by their coordinates in a high-dimensional feature space, which we 
construct using a relatively simple text embedding model (Latent Dirichlet 
Allocation;~\citealp{BleiEtal03}) trained on overlapping sliding windows of 
text from the lectures' transcripts (see \topicModelMethods). In comparing our 
approach to various alternative methods of characterizing the lectures' and 
questions' contents, we found that LDA embeddings exhibited a number of 
desirable properties that both simpler and more complex alternatives did not. 
For example, we found that analogous embeddings derived from more modern and 
complex models trained on much larger and more diverse text corpora (e.g., 
BERT,~\citealp{DevlEtal18}; GPT,~\citealp{ViswEtal17}; 
LLaMa,~\citealp{TouvEtal23}) tended to afford poor resolution within the content 
space spanned by individual course lectures and quiz questions. To illustrate 
this contrast, in Supplementary Figure~\ref{fig:compare-bert} we display a 
subset of visualizations from the main text based on our LDA embeddings (left 
column) alongside analogous visualizations based instead on BERT embeddings 
(right column).

>>>>

In our study, we relate the conceptual contents of course lectures and quiz 
questions by their coordinates in a high-dimensional feature space, which we 
construct using a relatively simple text embedding model (Latent Dirichlet 
Allocation;~\citealp{BleiEtal03}) trained on overlapping sliding windows of 
text from the lectures' transcripts (see \topicModelMethods). In comparing our 
approach to various alternative methods of characterizing the lectures' and 
questions' contents, we found that embeddings derived from our corpus-specific 
LDA model exhibited a number of desirable properties that both simpler and more 
complex alternatives did not. For example, we found that analogous embeddings 
derived from more modern and complex models trained on much larger and more 
diverse text corpora (e.g., BERT,~\citealp{DevlEtal18}; 
GPT,~\citealp{ViswEtal17}; LLaMa,~\citealp{TouvEtal23}) tended to afford poor 
resolution within the content space spanned by individual course lectures and 
quiz questions. To illustrate this contrast, in Supplementary 
Figure~\ref{fig:compare-bert} we display a subset of visualizations from the 
main text based on our LDA embeddings (left column) alongside analogous 
visualizations based instead on BERT embeddings (right column).
